% This is samplepaper.tex, a sample chapter demonstrating the
% LLNCS macro package for Springer Computer Science proceedings;
% Version 2.20 of 2017/10/04
%
\documentclass[runningheads]{llncs}
%
\usepackage{graphicx}
% Used for displaying a sample figure. If possible, figure files should
% be included in EPS format.
\usepackage{tikz}
\usetikzlibrary{arrows}
\usepackage{verbatim}
%\usepackage{amsmath}
%\usepackage{amssymb}
%\usepackage{graphicx}
%\usepackage[all]{xy}
\usepackage{array}
\usepackage{enumitem}
%\usepackage{cite}
\usepackage{natbib}
% If you use the hyperref package, please uncomment the following line
% to display URLs in blue roman font according to Springer's eBook style:
% \renewcommand\UrlFont{\color{blue}\rmfamily}
\usepackage[breaklinks=true]{hyperref}
\usepackage{breakcites}
\renewcommand\UrlFont{\color{blue}\rmfamily}

\begin{document}
%
\title{Understanding Program State Through Attestation}
%
%\titlerunning{Abbreviated paper title}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{Yan Shoshitaishvili\inst{1} \and Perry Alexander\inst{2}}
%
\authorrunning{Y. Shoshitaishvili and P. Alexander}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{Arizona State University \\
  Tempe, AZ \\
  \email{yans@asu.edu}
  \and
  Institute for Information Sciences \\ The
  University of Kansas \\ Lawrence, KS 66045 \\
  \email{palexand@ku.edu}}
%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
Runtime remote attestation is gathering verifiable evidence of program state
under constrained disclosure with the intent of establishing trust. A relying
party requests evidence from a remote target that responds by executing an
attestation protocol to gather and sign evidence and meta-evidence.  Designing
such attestation protocols requires identifying measurement points in program
state that provide meaningful approximations while respecting constrained
disclosure.  Finding such observation points is complicated by state complexity,
difficulty mapping execution state to behavior, and unavailability of source
code.  Static and dynamic analysis techniques including fuzzing, abstract
interpretation, and decompilation prove effective in approximating system state
for testing and debugging.  We propose repurposing such techniques to discover
measurements of user-space programs that soundly approximate runtime behavior.

\keywords{remote attestation \and static analysis \and runtime measurement \and program understanding.}
\end{abstract}
%
%
%

%%\section{Introduction}

\citet{Martin:08:The-ten-page-in} defines trust as exhibiting unambiguous
identification, unhindered operation, and direct observation of good behavior or
indirect observation by a trusted third party. Establishing trust in a remote
peer is a challenging problem requiring approximating program behavior by
measuring its state and appraising resulting evidence. One technique for
establishing trust in a target is semantic remote
attestation~\citep{Haldar:04:Semantic-Remote,Coker::Principles-of-R}.  Shown in
Figure~\ref{fig:remote-attestation} a \emph{relying party} ($RP$) sends an
attestation request ($r:(R,n,a)$) to a \emph{target} ($T$) where a protocol
gathers evidence and meta-evidence ($e:(E,n)$) that can be \emph{appraised} to
determine trust.

\begin{figure}[hbtp]
  \centering
  \begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node distance=2.0cm,
    thick,main node/.style={rectangle,fill=blue!20,draw,
      font=\sffamily,minimum height=20mm,minimum width=10mm},
    io node/.style={rectangle,
      font=\sffamily,minimum height=5mm,minimum width=10mm}]
    
    \node[main node] (NM) {$RP$};
    \node[main node] (NME) [node distance=5.0cm, right of=NM] {$T$};
%%    \node[io node] (IN) [above of=NM] {$R$};
%%    \node[io node] (OUT) [below of=NM] {$\mathtt{Prop}$};
    
    \path[every node/.style={font=\sffamily\small, fill=white,inner sep=1pt}]
    (NM) edge [bend left=30] node[above=1mm] {$\{(R,n,a)\}_{RP^{-1}}$} (NME)
    (NME) edge [bend left=30] node[below=1mm] {$\{(E,n)\}_{T^{-1}}$} (NM)
  %  (IN) edge (NM)
  %  (NM) edge (OUT)
    ;
  \end{tikzpicture}
  \caption{Remote attestation architecture
  showing a \emph{relying party} making an attestation request of a
  \emph{target}.}
  \label{fig:remote-attestation}
\end{figure}

\citet{Coker::Principles-of-R,Coker:08:Attestation:-Ev} define a remote
attestation model where a target chooses and executes an \emph{attestation
protocol} that gathers evidence and generates meta-evidence describing its
execution. The protocol sequences executing of attestation services including
measurement, signing, and making attestation requests of other systems.  These
protocols are executed by one or more \emph{attestation manager(s)} associated
with both the relying party and target.  Any attestation manager may choose not
to execute a protocol if its privacy policy would be violated.

Working with MITRE, JHUAPL, and NSA, the University of Kansas has developed the
Copland~\citep{Ramsdell:2019aa} attestation protocol language and the MAESTRO
framework~\citep{petz2022innovations,
Petz:2024:Verified-Configuration-and-Deployment-Paper} for synthesizing remote
attestation systems. Copland is a formally specified language that sequences the
execution of attestation services providers (ASPs) that gather measurement
information, perform cryptographic operations to generate meta-evidence, and
send requests to remote attestation managers. Using the Copland semantics,
verified Coq definitions of attestation, and synthesis of Coq to
CakeML~\citep{Kumar:2014:CVI:2535838.2535841} we have a formally verified
attestation manager that guarantees that good ASPs are executed in the right
order and at the right place.

MAESTRO configures the verified attestation manager for specific attestation
applications.  Given one or more Copland protocol, MAESTRO synthesizes manifests
and configuration files for a collection of attestation managers. MAESTRO tools
are formally verified and synthesize from Coq specifications to ensure
correctness of generate attestation systems.  If we have a performance envelope
defining good software execution, we can design Copland protocols and ASPs to
gather and appraise evidence indicating  when execution is in that envelop.

Difficulty finding the good performance envelop for arbitrary software systems
limits attestation's uptake.  Attestation targets are arbitrary software systems
whose behavior must be learned before writing a protocol.  Manually
understanding these systems is often infeasible due to lack of source code,
program complexity, and emergent behaviors.  Successful systems are bespoke
designs for individual systems that have limited potential for reuse.  For
Copland, MAESTRO and other attestation systems to succeed, finding good
performance envelopes and measurements that produce evidence for them must be
automated.

Arizona State University has developed
techniques to fuzz~\citep{trickel2022toss,salls2020exploring,peng2018t}, symbolically execute~\citep{stephens2016driller,shoshitaishvili2016sok}, and statically analyze~\citep{das2022hybrid,vadayath2022arbiter} complex user space programs and their components, including complex subsystems such as the dynamic allocator (``heap'')~\citep{eckert2018heaphopper} and cross-runtime interface bindings~\citep{dinh2021favocado}.
These techniques prove useful for automatically discovering critical data structures useful for attack and defense.

We propose a spectrum of techniques to automatically discover and implement
high-quality measurement for use in downstream attestation.  When source code is
available (increasing the efficacy of static analysis by granting it accurate
variable type information), we will use static analysis to identify memory
regions whose measurements will be particularly informative.  When source (and
variable type information) is not available, we will leverage our decompilation
techniques to find system variables whose values prove useful for appraisal, and
leverage fuzzing to identify unusual states.  Finally, we propose generating
ASPs and instrumenting code during compilation (when source is available) or
retrofitting it into the already-compiled binary using our binary modification
techniques~\citep{wang2017ramblr} (when source is not available) to provide
measurement access to critical structures.

We will use Fuzzing to generate and identify ``weird'' memory states.  Once such
states are identified, by analyzing the constraints that must be satisfied to
trigger them, we will develop measurement and appraisal routines to detect these
memory states.  Because we understand how fuzzing is performed, and can leverage
symbolic execution to delve deeper into the functionality of the program, we can
understand what measurements tell us about observed memory.

Program decompilation identifies critical system variables for runtime
measurement.  Decompilation produces source with variables present.  Once
decompiled, we will perform source-level analysis to determine what variables
are valuable for runtime measurement.  Using type information, we can bound
memory values and identify good and bad variable values.  Using this
information, we can automatically synthesize measurement and appraisal routines
for variable memory locations.

We plan to use static analysis and compilation techniques to automatically
generate measurement and appraisal routines from OS and user-space software.
For example, we can utilize our dynamic allocator analysis
results~\citep{eckert2018heaphopper} to synthesize measurement routines for
allocator implementation to detect integrity violations due to other bugs in the
program.  We can carry out similar intervention at the boundary between
different program components by leveraging interface specifications combined
with advanced analysis~\citep{dinh2021favocado}.  We must take care to limit the
performance impact of measurement instrumentation.  This problem is an
interesting mirror to false positive reduction during static analysis.  In other
words, the measurement should be triggered as rarely as possible during runtime
while still providing detection guarantees.  By leveraging our system of
\emph{Iterative False Positive Reduction} using a combination of Abstract
Interpretation and Dynamic Symbolic Execution~\citep{vadayath2022arbiter}, we
can narrow down to increasingly reduced instrumentation sites (by disproving the
need for instrumentation at given points in the program), reducing runtime
overhead by trading it for analysis-time cost.  Even in cases where we cannot
disprove the need for instrumentation, we may be able to ``back-propagate''
certain constraints that, if satisfied, disprove the need for further
measurement at runtime.  By carefully crafting such runtime instrumentation and
correlating it to marked memory locations and events, we can essentially
synthesize ``tripwires' that, when not tripped, cause very little overhead.

We proposed that using static analysis to discover performance envelopes will
improve quality of attestation systems while decreasing their cost to deploy. We
propose to: (i) extend the LKIM~\citep{Loscocco:07:Linux-kernel-in} approach of
defining run-time measurements and policies for checking evidence to user-space
applications; (ii) use static analysis to discover measurement targets and
appraisal policies defining good behavior; (iii) extend the MAESTRO tool suite
to include static analysis in defining attestation systems. Our experience with
LKIM suggests that run-time state can be measured and appraised successfully.
Static analysis has proven successful in discovering substantial
security-related structures in software systems.  Brining these two technologies
together will result in dramatic improvements in the way we build attestation
systems, their effectiveness, and our ability to understand the performance
envelop of software systems.

% 
% 
% ---- Bibliography ----
%
% BibTeX users should specify bibliography style 'splncs04'.
% References will then be sorted and formatted in the correct style.
%
%\bibliographystyle{splncs04}
\bibliographystyle{splncsnat}
\bibliography{sldg}
%
\end{document}
